---
title: "Titanic Project"
output: html_document
date: "2024-05-18"
---

```{r}
# Loading and viewing titanic data

data <-  read.csv("train")

View(train)

```


1.) What ages are found on the titanic and what is the most common age found on board?
```{r}
age_data <- na.omit(train$Age)

hist(age_data, main = "Age distribution on Titanic", xlab = "Age", ylab = "Frequency")

age_counts <- table(age_data)

most_common_age <- names(which.max(age_counts))

print("Most common age found on board:")
print(most_common_age)
```


2.) What is the most common passenger class found on board the Titanic?
```{r}
class_counts <- table(train$Pclass)

most_common_class <- names(which.max(class_counts))

print("Most common passenger class on board:")
print(most_common_class)

pie(class_counts, main = "Passenger Class Distibution on Titanic", labels = c("1st Class", "2nd Class", "3rd Class"), col = c("blue", "green", "gray"))
```
2). How many of each passenger class survived?

```{r}

survivors_by_class <- table(train$Pclass, train$Survived)

survivors_by_class <- survivors_by_class[, 2]  

print("Number of survivors from each passenger class:")
print(survivors_by_class)
```



3.) What gender was found most on board?
```{r}
sex_counts <- table(train$Sex)

most_common_sex <- names(which.max(sex_counts))

print("Most common sex found on board:")
print(most_common_sex)

barplot(sex_counts, main = "Sex distribution on Titanic", xlab = "Sex", ylab = "Frequency", col = "lightgreen")
```


4.) How many males compared to female were found on board the titanic? How many of each gender survived?

```{r}

num_males <- sum(train$Sex == "male")
num_females <- sum(train$Sex == "female")


num_male_survived <- sum(train$Sex == "male" & train$Survived == 1)
num_female_survived <- sum(train$Sex == "female" & train$Survived == 1)


print("Number of males:")
print(num_males)
print("Number of females:")
print(num_females)
print("Number of male survivors:")
print(num_male_survived)
print("Number of female survivors:")
print(num_female_survived)
```




5.)  How many families or siblings were found on board?
```{r}
total_siblings_spouses <- sum(train$SibSp)

print("Total number of siblings/spouses on board:")
print(total_siblings_spouses)
```

6.)How many passengers were on the Titanic total? How many of those passengers survived?
```{r}
total_passengers <- nrow(train)

survived_passengers <- sum(train$Survived == 1, na.rm = TRUE)

print("Total number of passengers on board:")
print(total_passengers)
print("Number of passengers who survived:")
print(survived_passengers)

survived_counts <- table(train$Survived)
barplot(survived_counts, main = "Survivors on Titanic", xlab = "Survived", ylab = "Frequency", names.arg = c("No", "Yes"), col = "lightblue")

```

7.) Which class survived the most?
```{r}
survivors_by_class <- table(train$Pclass, train$Survived)

most_survived_class <- rownames(survivors_by_class)[which.max(survivors_by_class[,2])]

print("Passenger class with the most survivors:")
print(most_survived_class)

```



8.) Of the survived passengers, were the majority male or female?
```{r}
survivors_by_sex <- table(train$Sex, train$Survived)

most_survived_sex <- rownames(survivors_by_sex)[which.max(survivors_by_sex[,2])]

print("Sex with the most survivors:")
print(most_survived_sex)

barplot(survivors_by_sex, beside = TRUE, 
        main = "Survivors and Non-Survivors by Sex",
        xlab = "Sex", ylab = "Frequency",
        legend = c("Not Survived", "Survived"),
        col = c("red", "green"),
        ylim = c(0, max(survivors_by_sex) * 1.2), 
        args.legend = list(x = "topright", bty = "n"),
        names.arg = c("Female", "Male"))

```



9.)  How many spouses/siblings made it out together?
```{r}
train$HasSpouseSibling <- ifelse(train$SibSp > 0, "Yes", "No")

survivors_with_spouse_sibling <- sum(train$Survived == 1 & train$HasSpouseSibling == "Yes")

print("Number of survivors among passengers with a spouse or a sibling on board:")
print(survivors_with_spouse_sibling)


```



10.) What age groups had the most survivors? Which age group did not?
```{r}
age_groups <- cut(train$Age, breaks = c(0, 18, 30, 50, Inf), labels = c("0-18", "19-30", "31-50", "51+"))

survivors_by_age_group <- table(age_groups, train$Survived)

most_survived_age_group <- rownames(survivors_by_age_group)[which.max(survivors_by_age_group[,2])]
most_non_survived_age_group <- rownames(survivors_by_age_group)[which.max(survivors_by_age_group[,1])]

print("Age group with the most survivors:")
print(most_survived_age_group)
print("Age group with the most non-survivors:")
print(most_non_survived_age_group)


```

11.) What passengers paid the highest fare? Did they survive?
```{r}
highest_fare <- max(train$Fare, na.rm = TRUE)
                    
passengers_highest_fare <- train[train$Fare == highest_fare, ]

print("Passengers who paid the highest fare and their survival status:")
print(passengers_highest_fare[, c("Name", "Survived")])
```

12.)  How many children were on board and how many survived?
```{r}
children <- train[train$Age <= 18, ]

total_children <- nrow(children)

children_survived <- sum(children$Survived == 1, na.rm = TRUE)


print("Total number of children on board:")
print(total_children)
print("Number of children who survived:")
print(children_survived)

```

`

